---
title: "Introduction to Linear Regression"
subtitle: "Data Science for Biologists, Spring 2020"
author: "Nicholas Paterna"
output: 
  html_document:
    highlight: pygments
    theme: readable
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
```

## Instructions

Standard grading criteria apply, except there is no "answer style" - just write out answers normally! Note the following:

+ $\alpha = 0.05$
+ When adding regression lines as plot subtitles, don't worry about writing $\epsilon$
+ Do not remove insignificant predictors from your regression line formula when reporting it (this is just a question I get a lot)

This assignment will use an external dataset from a field experiment studying the diversity of [Chinese Rowan](https://en.wikipedia.org/wiki/Rowan) trees. Researchers randomly sampled and recorded characteristics of leaves from three different species in the *Sorbus* genus. They recorded the following measurements for each tree (don't worry about units)

1. `species`: the species of tree
2. `altitude`: the altitude of the tree
3. `respiratory_rate`: average respiratory rate across a random sample of leaves from that tree
4. `leaf_len`: average leaf length across a random sample of leaves from that tree
5. `birds_nesting`: whether birds were actively nesting in the tree

For this assignment, you will examine how various predictors may explain variation in *respiratory rate.*

```{r}
# rowan trees, no relation :)
rowan <- read_csv("https://raw.githubusercontent.com/sjspielman/datascience_for_biologists/master/data/rowan_trees.csv")
dplyr::glimpse(rowan)
```

### Question 1

> Background for this completing question is under the header "Simple linear regression: Single numeric predictor" in the linear regression tutorial


Construct a linear model that examines how *altitude* might explain variation in *respiratory rate*. Take the following steps to do so (code as you go through the steps!)

Make a quick scatterplot to make sure the "linear relationship" assumption is met for this data. Be sure to have your predictor and response on the correct axis!:

```{r}
### figure to check linearity goes here.
### no need for fancy, can keep labels as is, etc.
ggplot(rowan, aes(x = altitude, y = respiratory_rate)) +
  geom_point() +
  labs(x = "Altitude", y = "Respiratory Rate")
```


Now that you have confirmed the relationship is linear (hint: it is linear), build your linear model. *BEFORE you examine its output*, evaluate whether the model residuals were normally distributed:

```{r}
### build model and check normal residuals
### do not look at the model summary yet!
fit <- lm(formula = respiratory_rate ~ altitude, data = rowan)
  #Always use y ~ x for formula!!!

augment(fit, rowan) -> augmented_fit
  #We use this function from broom to create a tibble with the residuals

qqnorm(augmented_fit$.resid, pch = 19)
qqline(augmented_fit$.resid, col = "dodgerblue")
  #Here the pch is the spacing of the points, we are plotting the residual values from the augmented fit. 
 #Using base R to create a linear regression with the residual values vs. the experimental observations.

```


Now that you have confirmed the residuals are roughly normally distributed (hint: they are), examine the output from your linear model. In the space below the code chunk, discuss in *bullet form* (1-2 sentences each): a) Provide an interpretation of the intercept, b) Provide an interpretation of the altitude coefficient, c) Provide an interpretation of the $R^2$ value (those dollar signs signify "math mode" - see the HTML output!), and finally d) Conclude whether altitude is a strong or weak predictor of respiratory rate, consider "biological significance" (effect size!) as well as statistical significance.

```{r}
## examine model output here
summary(fit)
```

+ **Intercept Interpretation**
  + We expect that a tree's *respiratory rate* will be 8.386e-03 when the **altitude is 0**. However, p is not significant at p = 0.622 so we actually have no evidence that intercept differs from 0.

+ **Altitude Coefficient Interpretation**
  + For every unit increase in altitude, we expect respiratory rate increase by 3.025e-04, this is highly significant because our p-value is so small for this measurment (p = 2.66e-14).

+ **$R^2$ Interpretation**
  + $R^2 = 0.1743$ which means that approx. 17% of variation in respiratory rate is explained by our model, which included only altitude predictir. This is highly significant as the corresponding p-value is so small (p = 2.66e-14).
  
  + From this, approximately 83% of these respiratory rate cannot be explained by this model. 
  
+ **Model Conclusion**
  + Y = 3.025e-04X + 8.386e-03

Finally, make a stylish scatterplot of your findings. Your scatterplot should:

+ Use your favorite ggplot theme and colors (it's allowed to like the default!)
+ Clearly show the regression line and its 95% confidence interval
+ Include a meaningful title, along with a subtitle that is the fitted model itself, as well as other nice labels
+ Include a well-placed annotation that gives the model's $R^2$

```{r}
### stylish plot goes here
ggplot(rowan, aes(x = altitude, y = respiratory_rate)) +
  geom_point() +
  labs(title = "Regression of Altitude Explaining Resp Rate",
       subtitle = "Y = 3.025e-04X + 8.386e-03",
       x = "Altitude", y = "Respiratory Rate") +
  geom_smooth(method = "lm", color = "hotpink") +
  #Default confidence interval (shaded area) is 95%
  annotate("text", label = "R^2 == 0.1743",
           x = 200, y = 0.3,
           size = 5,
           parse = TRUE) +
  theme_minimal()

```


### Question 2

> Background for this completing question is under the header "Simple ANOVA: Single categorical predictor" in the linear regression tutorial

Construct a linear model that examines how *species* might explain variation in *respiratory rate*. Take the following steps to do so (code as you go through the steps!)

Make a quick plot (use `geom_point()`, seriously, not even a jitter!!) to make sure the "equal variance" assumption is met for this data:

```{r}
### figure to check assumption goes here.
### no need for fancy, can keep labels as is, etc.
ggplot(rowan, aes(x = species, y = respiratory_rate)) +
  geom_point()
#This essentially shows us the variation between the resp rate over the three species, it is clear that they are fairly similar.

```


Now that you have confirmed the variance is equal across groups (hint: it is), build your linear model. *BEFORE you examine its output*, evaluate whether the model residuals were normally distributed:

```{r}
### build model and check normal residuals
### do not look at the model summary yet!

fit_2 <- lm(formula = respiratory_rate ~ species, data = rowan)
  #Always use y ~ x for formula!!!

augment(fit, rowan) -> augmented_fit2
  #We use this function from broom to create a tibble with the residuals

qqnorm(augmented_fit2$.resid, pch = 19)
qqline(augmented_fit2$.resid, col = "goldenrod")
  #Here the pch is the spacing of the points, we are plotting the residual values from the augmented fit. 
 #Using base R to create a linear regression with the residual values vs. the experimental observations.

```

Now that you have confirmed the residuals are roughly normally distributed (hint: they are), examine the output from your linear model. In the space below the code chunk, discuss in *bullet form* (1-2 sentences each): a) Provide an interpretation of the intercept, b) Provide an interpretation of the species coefficient, c) Provide an interpretation of the $R^2$ value, and finally d) Conclude whether species is a strong or weak predictor of respiratory rate, consider "biological significance" (effect size!) as well as statistical significance.

```{r}
## examine model output here
summary(fit_2)

unique(rowan$species)
#This shows us the unique variables in the species col so we can determine which of these was the "baseline" for our model

```

+ **Intercept Interpretation**
  + For microphylla trees we expect a mean respiratory rate of 0.148680, this is highly significant as our p-value is so small (p < 2e-16).
  
+ **Species Coefficient Interpretation**
  + We expect on average, *oligodonta* trees have a respiratory rate that is 0.018830 lower than microphylla. However, this is not significant due to the high p-value (p = 0.0995).
    + We expect on average, *argentiana* trees have a respiratory rate that is 0.006180 lower than microphylla. However, this is not significant due to the high p-value (p = 0.5880).

+ **$R^2$ Interpretation**
  + Our $R^2 = 0.002792$ which is not NOT SIGNIFICANT because the p-value is so large (p = 0.2437). No evidence that any variation is explained, 100% of variation in respiratory rate is unexplained by model.

+ **Model Conclusion**
  + Entirely uninformative model, species does NOT predict or explain the respiratory rate of the trees.
  
  + Y = -0.018830X1 + -0.006180X2 + 0.148680

Finally, make a stylish figure of your findings, choosing your own geom!

+ Use your favorite ggplot theme and colors (it's allowed to like the default!)
+ If your geom does not already show the center of each group (i.e. like a boxplot), be sure to add the means in with `stat_summary()`
+ Include a meaningful title, along with a subtitle that is the fitted model itself, as well as other nice labels
+ Include a well-placed annotation that gives the model's $R^2$

```{r}
### stylish plot goes here
ggplot(rowan, aes(x = species, y = respiratory_rate)) +
  geom_jitter(width = 0.15) +
  stat_summary(color = "yellowgreen")
  #Adds a point into the plot that shows the std error for all of the points

```


### Question 3

> Background for this completing question is under the header "LM with numeric and categorical predictors" in the linear regression tutorial

Construct a linear model that examines how BOTH *species* and *leaf_len* as independent effects might explain variation in *respiratory rate*. Again, take the following steps one by one:

Since we already checked assumptions for `species` in the last question, make an appropriate plot to check the linearity assumption for `leaf_len`:

```{r}
### figure to check assumption goes here.
### no need for fancy, can keep labels as is, etc.
ggplot(rowan, aes(x = leaf_len, y = respiratory_rate)) +
  geom_point()

```

Build your linear model, and evaluate whether the model residuals were normally distributed:

```{r}
### build model and check normal residuals
### do not look at the model summary yet!
fit_3 <- lm(respiratory_rate ~ species + leaf_len, data = rowan)

augment(fit_3, rowan) -> augmented_fit3
  #We use this function from broom to create a tibble with the residuals

qqnorm(augmented_fit3$.resid, pch = 19)
qqline(augmented_fit3$.resid, col = "forestgreen")
  #Here the pch is the spacing of the points, we are plotting the residual values from the augmented fit. 
 #Using base R to create a linear regression with the residual values vs. the experimental observations.

```

Now that you have confirmed the residuals are roughly normally distributed (hint: they are), examine the output from your linear model. In the space below the code chunk, discuss in *bullet form* (1-2 sentences each): a) Provide an interpretation of the intercept, b) Provide an interpretation of the `species` coefficient, c) Provide an interpretation of the `leaf_len` coefficient d) Provide an interpretation of the $R^2$ value, and finally e) Conclude whether species is a strong or weak predictor of respiratory rate, consider "biological significance" (effect size!) as well as statistical significance.

```{r}
## examine model output here
summary(fit_3)

```

+ **Intercept Interpretation**
  + For microphylla trees we expect a respiratory rate of 0.176 when **leaf length is 0**, this is highly significant (p < 2e-16). *However, this is biologically unrealistic.*
  
+ **Species Coefficient Interpretation**
  + When controlling for leaf length, coefficients are the change from baseline species

  + Controlling for leaf length, *oligodonta* have an average respiratory rate of 0.003898 lower than microphylla. But this is not significant (p = 0.0375), so effectively there is no difference between these species.
  
  + Controlling for leaf length, *argentiana* have an average respiratory rate of 0.050740 higher than microphylla. But this is not significant (p = 0.0864), so effectively there is no difference between these species.

+ **$R^2$ Interpretation**
  + Our $R^2 = 0.01398$ which is not NOT SIGNIFICANT because the p-value is large (p = 0.06687). No evidence that any variation is explained, 100% of variation in respiratory rate is unexplained by model.

+ **Model Conclusion**
  + Entirely uninformative model, leaf length does NOT predict or explain the respiratory rate of the trees.

Finally, make a stylish scatterplot of your findings:

+ Use your favorite ggplot theme and colors (it's allowed to like the default!)
+ Make sure to show a regression lines for EACH species. **NOTICE in theses lines** how they are consistent with your conclusions about species being a significant predictor. You do not have to write anything, just notice!
+ Include a meaningful title, along with a subtitle that is the fitted model itself, as well as other nice labels
+ Include a well-placed annotation that gives the model's $R^2$

```{r}
### stylish plot goes here
ggplot(rowan, aes(x = leaf_len, y = respiratory_rate, color = species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Leaf Length", y = "Respiratory Rate", color = "Species") +
  theme_minimal()
#The overlapping of the conf intervals shows that this model is poor.
```

