---
title: "#TidyTuesday Submission"
subtitle: "Data Science for Biologists, Spring 2020"
author: "Nicholas Paterna"
date: "March 15, 2020"
output:
  html_document:
    highlight: tango
    theme: yeti
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(tidyverse)
library(tidylog)
theme_set(theme_minimal())

tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')


```

### TidyTuesday Week 11
Hi, this is my fourth time attempting to make a figure for #TidyTuesday! I am a student at Rowan University in Biochemistry with **_a smidge_** of experience in coding thanks to [Dr. Stephanie Spielman](https://twitter.com/stephspiel). R is my first language and I am very happy to be learning it, I hope to work in Bioinformatics one day possibly Medical Genomics. I am open to criticism or helpful tips in trimming my code!!

```{r Variable Coding}

tidy_path <- "/Users/nickpaterna/datascience/TidyTuesday/Week_11/"
#This variable will tell the figure where to save when I apply ggsave!

figure_name <- "tidytues_week11.png"
#Rather than re-typing this into ggsave I can just use this!

fav_fills <- c("aquamarine", "azure3")
#Assigned fills to the figure.

position_labels <- c("Center", "Left-Wing", "Right-Wing")
#New labels for each of the positions.
```

## Avoid Hard-coding!
One of the most important things that my professor stresses is a complete aversion to HARD-CODING, or explicitly re-typing code in a large script. It might not make a lot of sense to apply these practices on a small assignment but the more often I apply it the easier to apply it is in the future! (Also the second visible code chunk gives my submission some more flavor!)


```{r TT_Week11, fig.height=4, fig.width=6}

tuition_cost %>%
  select(degree_length, in_state_total, out_of_state_total) %>%
  #Isolating the cols of interest to me.
  filter(degree_length %in% c("2 Year", "4 Year")) %>%
  pivot_longer(in_state_total:out_of_state_total,
               names_to = "resident",
               values_to = "total_cost") %>%
  #Using this function to combine the cols I want.
  mutate(resident = str_replace_all(resident,
                                    c("in_state_total" = "in state",
                                    "out_of_state_total" = "out of state"))) %>%
  mutate(resident = str_to_title(resident)) -> tidy_tuition

head(tidy_tuition, 10)
#Displaying the tidy data to show changes.
#Renaming the values in the cols using stringr & saving the tidy data as a variable.
  
tidy_tuition %>%
  ggplot(aes(x = total_cost, fill = resident)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values = fav_fills) +
  labs(x = "Total Cost", y = "Density", fill = "Residence") -> tuition_plot
  
ggsave(paste0(tidy_path, figure_name), tuition_plot, height = 4, width = 6)
tuition_plot
```

This week I wanted to use the skills we learned this past week in creating a plot that reflected the total cost (room & board + tuition) for students based on residency. I know that when I graduated high school cost was the most important factor when deciding where to go!

Here I tidied the data using the `pivot_longer()` function which I **LOVE** and used my new [stringr](https://stringr.tidyverse.org/reference/index.html) skills to tidy the output from this. I wanted to go a little further by faceting my figure by degree length or type but the output wasn't satisfying or necessarily informative so I stuck with what I had. I also avoided hard-coding again which is always beneficial! Check out my [GitHub](https://github.com/npaterna/datascience/tree/master/TidyTuesday) for past and future submissions!!