---
title: "Clustering and PCA"
subtitle: "Data Science for Biologists, Spring 2020"
author: "Nicholas Paterna"
output: 
  html_document:
    highlight: zenburn
    theme: yeti
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(tidyverse)
theme_set(theme_minimal())

set.seed(1227)
# My brother's birthday!
```

## Instructions

Standard grading criteria apply. And...

+ Make sure you set your seed
+ Proofread to submit YOUR OWN WORDS!!
+ Make sure all bullets properly render in the knitted HTML. **YOU WILL LOSE POINTS FOR BULLETED LISTS THAT DO NOT TURN OUT PROPERLY IN THE HTML!!** 
+ Remember that `#` do NOT make comments in markdown. They make HEADERS!! You need to use HTML comments to have comments in your text that don't appear in the knitted HTML. For example, the line below does not appear in html.

<!-- This is an actual comment in markdown. -->

## The Data

We will be using a dataset of olives collected from various regions of Italy. The data gives the percentage of different fatty acids that comprise the oil in the olives. [Source](https://rdrr.io/cran/dslabs/man/olive.html)

```{r, collapse=T}
olives <- read_csv("https://raw.githubusercontent.com/sjspielman/datascience_for_biologists/master/data/olives.csv")
dplyr::glimpse(olives)
```


## Part 1: Clustering

Perform TWO k-means cluster analysis, once with **k=3** and once with **k=5**. 
```{r}
## Perform two clustering analyses here with k=3 and k=5, separately.
olives %>%
  select(-region, -area) %>%
  kmeans(3) -> olives_k3

olives %>%
  select(-region, -area) %>%
  kmeans(5) -> olives_k5

```


For each analysis, make a SINGLE FIGURE for each clustering analysis (two figures total). Choose your own adventure for each figure! You should explain per figure: a) Why did you decide to visualize the clustering using this particular figure? (~1 sentence) and b) What insights (if any!) does this figure provide (1-2 sentences)? Answer in the bullets below the given figure. 

**You should not make two identical figures - they need to be _somewhat_ different.** For example, two barplots showing the relationship between clusters and regions is a no-go. Two scatterplots but with different x-axis and y-axis variables on the axes is ok! Two boxplots (violin, den sity, histogram, sina, jitter...) but with a different focal variable is ok! Either way, your figure MUST show the clusters.

```{r}
## Figure for k=3 including all associated wrangling
## We are using k = 3 for this cluster

olives %>%
  mutate(k3_clusters = factor(olives_k3$cluster)) %>%
  ggplot(aes(x = k3_clusters, y = stearic, fill = k3_clusters)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Spectral") +
  labs(x = "K-Means Clusters", y = "Stearic Acid") +
  theme(legend.position = "NULL")

```

+ **Answer: Why did you choose this figure?**
  + I wanted to investigate how the clusters grouped data by stearic acid content.

+ **Answer: What trends/insights (if any)  does the figure reveal, as they relate to the clusters?**
  + This figure does not necessarily provide insight into the clusters created but it does show that stearic acid content is similar among most of the olives evaluated.


```{r}
## Figure for k=5 including all associated wrangling
## We are using k = 5 for this cluster

olives %>%
  mutate(k5_clusters = factor(olives_k5$cluster)) %>%
  ggplot(aes(x = palmitic, y = linoleic, color = k5_clusters)) +
  geom_point() + stat_ellipse() +
  scale_fill_brewer(palette = "Spectral") +
  labs(x = "Palmitic Acid", y = "Linoleic Acid", color = "K-Means Cluster")

```

+ **Answer: Why did you choose this figure?**
  + I chose to compare two of the acids present in the olives evaluated and how they were clustered.

+ **Answer: What trends/insights (if any) does the figure reveal, as they relate to the clusters?**
  + There was more variation between the points not only because they are grouped into *more* clusters but also the evaluated points drift further from the centroids.
  
  + Therefore, I believe that these acids either have no relation or that 5 clusters are insufficient for this data.


## Part Two: PCA

Perform a principal components analysis on this dataset, and create TWO figures from your PCA. Each figure should have a brief explanation as a bullet below it. For each figure, your should be labeled to include information about how much variation the given PC explains in the data.

```{r}
## Code to perform PCA goes here
olives %>%
  select(-region, -area) %>%
  scale() %>%
  prcomp() -> olives_pca

```

1. But first, use code to answer this question! Intepret how variables in the initial dataset **load** on the PCs by examining, you guessed it, the loadings. In the bullets below, state which SINGLE variable load most strongly on PC1 and PC2 each, and which SINGLE variable loads most weakly on PC1 and PC2.

```{r}
## Code to examine loadings goes here
olives_pca$rotation

```

+ **Strongest PC1 loading:**
  + PC1 has the *strongest* overall relationship with **oleic acid**.

+ **Strongest PC2 loading:**
  + PC2 has the *strongest* overall relationship with **linolenic acid**.

+ **Weakest PC1 loading:**
  + PC1 has the *weakest* overall relationship with **stearic acid**.

+ **Weakest PC2 loading:**
  + PC2 has the *weakest* overall relationship with **palmitic acid**.

2. Plot PC2 against PC1 (PC2 is on Y and PC1 is on X) as a scatterplot, where you use color to distinguish among **olive region** (NOT area!). Make sure to show ellipses for each grouping. Do NOT include loading arrows. In the bullet below, _explain in 2-3 sentences_ if and how either PC1 or PC2 discriminates olive region. You may write a bulleted list *as long as it renders properly.*

```{r}
## Code to make this figure goes here
as_tibble(olives_pca$x) %>%
  mutate(region = olives$region) %>%
  ggplot(aes(x = PC1, y = PC2, color = region)) +
  geom_point() + stat_ellipse() +
  #stat_ellipse() easily visualizes the clustering by region
  scale_color_brewer(palette = "Dark2") +
  labs(color = "Region")
  
```

+ **Explanation for Figure:**
  + This figure shows how the PCA method successfully groups points by their region with few outliers.
  
  + Therefore, we can appropriately identify the reigon an olive originates from with **some** certainty from this figure.

2. Plot PC2 against PC1 as a scatterplot, where you use color to distinguish among **olive area** (the other categorical column!). Make sure to show ellipses for each grouping. Do NOT include loading arrows. In the bullet below, _explain in 2-3 sentences_ if and how either PC1 or PC2 discriminates olive region. You may write a bulleted list *as long as it renders properly.*

```{r}
## Code to make this figure goes here
as_tibble(olives_pca$x) %>%
  mutate(area = olives$area) %>%
  ggplot(aes(x = PC1, y = PC2, color = area)) +
  geom_point() + stat_ellipse() +
  #stat_ellipse() easily visualizes the clustering by region
  labs(color = "Area")
  
```

+ **Explanation for Figure:**
  + This figure shows how the PCA method successfully groups points by their region with few outliers.
  
  + Therefore, we can appropriately identify the area an olive originates from with **some** certainty from this figure, this is not as *clean* as the figure made from our regional data.

## BONUS QUESTION 1

Make a figure of the loadings on PC1 and PC2 in the space below. Make sure it is as legible as possible! No explanation needed, just make a plot!
```{r}
## Bonus loadings plot
```

## BONUS QUESTION 2

Perform a k-means clustering with *k=3* on the PCs to examine whether olive *regions* cluster together. **You must include a clear explanation (1-3 sentences) to receive ANY CREDIT AT ALL for this bonus.**
```{r}
## Bonus clustering
```
+ **Bonus explanation**
