---
title: "#TidyTuesday Submission"
subtitle: "Data Science for Biologists, Spring 2020"
author: "Nicholas Paterna"
date: "March 22, 2020"
output:
  html_document:
    highlight: tango
    theme: yeti
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
library(tidyverse)
library(tidylog)
library(patchwork)
theme_set(theme_minimal())

office_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-17/office_ratings.csv')

```

### TidyTuesday Week 12
Hi, this is my fifth time attempting to make a figure for #TidyTuesday! I am a student at Rowan University in Biochemistry with **_a smidge_** of experience in coding thanks to [Dr. Stephanie Spielman](https://twitter.com/stephspiel). R is my first language and I am very happy to be learning it, I hope to work in Bioinformatics one day possibly Medical Genomics. I am open to criticism or helpful tips in trimming my code!!

```{r Variable Coding}

tidy_path <- "/Users/nickpaterna/datascience/TidyTuesday/Week_12/"
#This variable will tell the figure where to save when I apply ggsave!

figure_name <- "tidytues_week12.png"
#Rather than re-typing this into ggsave I can just use this!
```

## Avoid Hard-coding!
One of the most important things that my professor stresses is a complete aversion to HARD-CODING, or explicitly re-typing code in a large script. It might not make a lot of sense to apply these practices on a small assignment but the more often I apply it the easier to apply it is in the future! (Also the second visible code chunk gives my submission some more flavor!)


```{r TT_Week11, fig.height=4, fig.width=6}

office_ratings %>%
  select(season, imdb_rating, total_votes) %>%
#Isolating columns of interest.  
  group_by(season) %>%
  summarize(avg_rating = mean(imdb_rating),
            tot_votes = sum(total_votes)) -> tidy_office
#Creates two new columns with total votes & average rating, summarized by season.

tidy_office
#Displaying the tidy data to show changes.
  
tidy_office %>%
  ggplot(aes(x = season, y = avg_rating, fill = season)) +
  geom_col(alpha = 0.7) +
  scale_fill_gradient(low = "#6ca66a", high ="#2f4f48") +
#I love this function, I use the ColorPicker on Google to find the Hex#s for the colors I want to see in my figure.  
  labs(x = "Season", y = "Average Rating") +
  theme(legend.position = "none") +
  ylim(0 ,10) -> half_officeplot

tidy_office %>%
  ggplot(aes(x = season, y = tot_votes, fill = season)) +
  geom_col(alpha = 0.7) +
  scale_fill_gradient(low = "#6ca66a", high ="#2f4f48") +
  labs(x = "Season", y = "Total Votes", fill = "Season") -> sechalf_officeplot

(half_officeplot + sechalf_officeplot) -> office_plot
#Using this awesome package called 'patchwork' I was able to make two separate plots, save them as variables, and then glue them together with a little electronic scrapbooking.

ggsave(paste0(tidy_path, figure_name), office_plot, height = 4, width = 6)
office_plot
```

I almost forgot to make my plot this week because I have been SO engrossed in Animal Crossing: New Horizon. I am not going to lie, I would love to see some AC themed datasets in future #TidyTuesday repos.

I have never seen The Office, I know shame on me, so I decided to go for simplicity with observations I could understand without utilizing the Shrute package. I couldn't decide between the average rating per season or the total number of votes so I decided to just throw both together! I used a package that we utilized in class called `patchwork` and it is, for lack of better word, delicious. I hope you enjoy my plot, you can find this and future entries on my [GitHub](https://github.com/npaterna/datascience/tree/master/TidyTuesday).